### 2017-04-20 CS798-ConvexityAndOptimization/assignments/assignment_03.tex
problem 2 solved

%Question 3
\section{}
\end{document}

>>>>

Let $A \in \{\pm 1\}^{m\times n}$. Suppose that the rank of $A$ is $r$. Let $a^1, \dots, a^r \in \R^m$ be a basis for the rowspace of $A$. Then for each row $i \in [m]$ of $A$ there exists $\lambda_1^i, \dots, \lambda^i_r$ such that the row vector can be expressed as $\sum_{j=1}^r \lambda_j^i a^j$. Therefore if we let $u_i^T = \begin{bmatrix} \lambda^i_1 &\dots &\lambda^i_r\end{bmatrix}$ and collect $U = \begin{bmatrix} u_1 &\dots & u_m \end{bmatrix}$, and we let $$V = \begin{bmatrix}
(a^1)^T \\ \dots \\ (a^r)^T
\end{bmatrix}$$ then we can write $A$ as 
$$A = U^T V.$$
Let $v_1, \dots, v_n$ be the columns of $V$. Since $A \in \{\pm 1\}^{m \times n}$, observe that $$|u^T_i v_j | = 1$$ for all $i \in [m], j \in [n]$. This fact will be useful later.
Let $K$ be the convex hull of $\{ \pm u_i : i \in [m]\}$. Then $K$ is a symmetric convex body in $\R^r$. By John's theorem there exists an ellipsoid $\cE$, centered at $0$, such that
$$ \frac{1}{\sqrt{r}} \cE \subseteq K \subseteq \cE.$$
There exists some positive definite matrix $H$ such that we can write $\cE = \{ x : x^TH^{-1}x \leq 1\}$. Then $H^\frac{1}{2} (\cE) = \cB$ where $\cB$ is the Euclidean norm unit ball. So $\sqrt{r}H^\frac{1}{2}(\cE) = \sqrt{r}\cB$. Let $\overline{U} = \sqrt{r}H^\frac{1}{2}U$, and let $\overline{V} = \frac{1}{\sqrt{r}} H^\frac{-1}{2}V$. Then
$$\overline{U}^T\overline{V} = U^TH^\frac{1}{2}H^\frac{-1}{2}V = U^TV = A$$
using that $H^\frac{1}{2}$ is symmetric.
For any matrix $M$ we use $col(M)$ to denote the set of column vectors of $M$. Since $col(U) \subseteq K$ and $$\sqrt{r}H^\frac{1}{2}(K) \subseteq \sqrt{r}H^\frac{1}{2} \cE = \sqrt{r} B$$
we have $||\overline{u}|| \leq \sqrt{r}$ for all $\overline{u} \in col(\overline{U}) = col(\sqrt{r}H^\frac{1}{2}U)$. Let $\overline{v} \in col(\overline{V})$. Then 
$$||\overline{v}||^2 = \overline{v}^T \overline{v} = \frac{1}{r}v^THv = \frac{1}{r} ||v||_H^2.$$
Thus
$$||\overline{v}|| = \frac{1}{\sqrt{r}} ||v||_H = ||v||_{\frac{1}{\sqrt{r}}H}.$$
Now the norm $||\cdot ||_{\frac{1}{\sqrt{r}}H}$ is dual to the norm $||\cdot||_{\frac{1}{\sqrt{r}}H^{-1}}$ so 
$$||\overline{v}|| = \sup \{ v^Tx : \frac{1}{\sqrt{r}} x^TH^{-1}x \leq 1\} = \sup_{x \in \frac{1}{\sqrt{r}}\cE} x^Tv.$$
Since $\frac{1}{\sqrt{r}}\cE \subseteq K$ we have
$$||\overline{v}|| \leq \sup_{x \in K} x^Tv.$$
Since $K$ is the convex hull of $\{\pm u : u \in col(U)\}$, for all $x \in K$ there exists $\lambda \in \R^{2r}$ such that $\lambda \geq 0$, $\sum_{i=1}^{2r} \lambda_i = 1$ and $\sum_{i=1}^r \lambda_i u_i - \sum_{i=r+1}^{2r} \lambda_i u_i = x$. But, since $|u^Tv| =1$ for all $u \in col(U)$, this expression is maximized when $\lambda_{r+1} = \dots = \lambda_{2r} = 0$ and in such a case 
$$\sum_{i=1}^r \lambda_i u_iv^T \leq \sum_{i=1}^r\lambda_i\cdot 1 = 1.$$
Hence
$$||\overline{v}|| \leq \sup_{x \in K} x^Tv \leq 1.$$
Therefore, for all $\overline{v} \in col(\overline{V})$, $||\overline{v}|| \leq 1\leq \sqrt{r}$, and for all $\overline{u} \in col(\overline{U})$, $||\overline{u}|| \leq \sqrt{r}$. Further
$$A = \overline{U}^T\overline{V}$$
%Question 3
Let $f(x) = \exp(-h(x))$ where $h$ is a differentiable function.
\subsection{a}
We want to show $f$ is log-concave (ie. $h$ is convex) if and only if $f''(x)f(x) \leq (f'(x))^2$. We compute the first and second derivatives of $f$. The first derivative is
$$f'(x) = \exp(-h(x))\cdot (-h'(x)).$$
The second derivative is 
$$f''(x) = \exp(-h(x)) \cdot (h'(x))^2 + \exp(-h(x))(-h''(x)) = \exp(-h(x))((h'(x))^2 - h''(x)).$$
So the left hand side is
$$f''(x)f(x) = \exp(-h(x))^2((h'(x))^2 - h''(x))$$
and the right hand side is
$$(f'(x))^2 = \exp(-h(x))^2(h'(x))^2.$$
So we have have $f''(x)f(x) \leq (f'(x))^2$ if and only if $h''(x) \geq 0$ for all $x$. But this second order condition is satisfied if and only if $h$ is convex, and therefore $f$ is log-concave.$\blacksquare$
\subsection{b}
Suppose $h$ is convex. Let $F(x) = \int_{-\infty}^x f(y) dy$. Let $x \in \R$ and suppose that $h'(x) < 0$. From the fundamental theorem of calculus, $F'(x) = f(x)$. Now let 
$$I(x) = -h'(x)F(x)$$.
Then $$I'(x) = -h''(x)F(x) -h'(x)F'(x) \leq -h'(x)f(x)$$
since $h''(x) \geq 0$ as $h$ is convex. Taking antidervatives we see
$$I(x) \leq f(x).$$
Thus
$$F(x) \leq \frac{f(x)}{-h'(x)}$$
since $h'(x) < 0$. So we have
$$\int_{-\infty}^xf(y)dy \leq \frac{\exp(-h(x))}{-h'(x)}$$
Now we show that $F$ is log-concave. We have
$$F'(x) = f(x)$$
and
$$F''(x) = -h'(x)f(x).$$
$$F''(x)F(x) = -h'(x)f(x) F(x)$$
and
$$(F'(x))^2 = f(x)^2.$$
If $h'(x) \geq 0$ then 
$$F''(x) F(x) \leq 0$$
and so $F''(x)F(x) \leq (F'(x))^2$ immediately. Now if $h'(x) < 0$ then we use the inequality we just demonstrated, so
$$F''(x) F(x) = -h'(x)f(x)F(x) \leq -h'(x)f(x) \cdot \frac{f(x)}{-h'(x)} = f(x)^2.$$
Therefore by part $3a$, $F(x)$ is log-concave. $\blacksquare$
\end{document}
